Dirac neutrino mass generation from Majorana messenger

### 2019-09-20 draft.tex
Updates from Overleaf

The results for $\Delta N_{\text{eff}}$ as a function of $M_{Z^{\prime}}/g^{\prime}$ are displayed in Fig.~\ref{fig:Neff}, where it can be observed that for small ratios $M_{Z'}/g'$  the variation in the number of relativistic degrees of freedom of the new species is large ($\Delta N_{\text{eff}}\gtrsim0.3$ corresponds to decoupling temperatures $T^{\nu_R}_{\text{dec}}\lesssim2$ GeV) and vice versa.
% The bound that come from Planck~\cite{Aghanim:2018eyx} is stronger than the bound of LEP~\cite{Heeck:2014zfa}. and even LHC, see below
It follows that for the $\operatorname{U}(1)_{G}$ model the Planck+BAO upper bound at $2\sigma$ (solid gray line) demands that $M_{Z^{\prime}}/g^{\prime} \gtrsim 40$ TeV while for the $\operatorname{U}(1)_{R}$ model a more stringent lower bound applies $M_{Z^{\prime}}/g^{\prime} \gtrsim 60$ TeV (these bounds become slightly weaker once the Planck+BAO+$H_0$ combination~\cite{Aghanim:2018eyx} is considered, but in such a case the BBN bound from the primordial abundances of light elements would rule~\cite{Pitrou:2018cgg}). 
%in order to satisfy the conditions of $N_{\text{eff}}$ for all the symmetries studied here.
In other words, bearing in mind that $M_{Z'}/g'=3v_S/2$ we have that the energy scale of the $\operatorname{U}(1)_{B-L}$ symmetry breaking must be at least $\sim26$ TeV.      
On the other hand, it is remarkable the fact the next generation of CMB experiments~\cite{Benson:2014qhw,Abitbol:2019nhf} has the potential to entirely probe all the not-so minimal models~\cite{Abazajian:2019oqj}.  
%Prospect on $N_{\text{eff}}$~\cite{Abazajian:2019oqj}

@@

\section*{Acknowledgments}
Work supported by Sostenibilidad-UdeA and the UdeA/CODI Grant
2017-16286, and by COLCIENCIAS through the Grant 111577657253. D.R
would also like to thank the IIP-Natal, for its support,
hospitality and the nice and stimulating atmosphere during the
Visiting Professors program where part of this work was completed.
\appendix
\section{$\operatorname{U}(1)_X$ anomaly cancellation }
We use $f$ ($f$) to denote the general $\operatorname{U}(1)_X$ generation-independent charge assignments of the field $f_R$ ($F_L$). The three linear anomalies in $\operatorname{U(1)_X}$~\cite{Campos:2017dgc}
\begin{align}
  \label{eq:anolin}
  \left[\mathrm{SU}(3)_{C}\right]^{2} \mathrm{U}(1)_{X} :& &
[3 u+3 d]-[3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{SU}(2)_{L}\right]^{2} \mathrm{U}(1)_{X} :&&
-[2 l+3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{U}(1)_{Y}\right]^{2} \mathrm{U}(1)_{X} :&&
{ \left[(-2)^{2} e+3\left(\tfrac{4}{3}\right)^{2} u+3\left(-\tfrac{2}{3}\right)^{2} d\right]-\left[2(-1)^{2} l+3 \cdot 2\left(\tfrac{1}{3}\right)^{2} q\right]}=&
 0,
  \end{align}
allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $N$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \left[\text{Grav}\right]^{2} \mathrm{U}(1)_{X}:\ \sum_{\alpha=1}^N n_{\alpha}+3 (e-2l)=&0\,, &  \left[\mathrm{U}(1)_{X}\right]^{3}:\ \sum_{\alpha=1}^N n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
We choose the solutions with $r\equiv e-2l$, such that
\begin{align}
  \label{eq:anolam}
  \sum_{\alpha=1}^{N} n_{\alpha}=&-3 r\,,   & \sum_{\alpha=1}^{N} n_{\alpha}^3=&-3 r^3\,.
\end{align}
The full set of anomaly free standard model $X$-charges in terms of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc} that we choose as $l$ and $r$, is just
\begin{align}
  u=&-r-\frac{4l}{3}\,,&d=&r+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&r+2l\,,&h=&-r-l\,.
  \label{Eq:SMCharges}
\end{align}
where the condition in the charged lepton Yukawa couplings have been used to fix $h$, and is automatically consistent with the conditions in the quark Yukawa couplings. By setting $l=0$ in the previous equations, we can define the Abelian symmetry in which only the right-handed charged fermions have non-vanishing $X$-charges as $\operatorname{U}(1)_R$. Then the general anomaly free two-parameter solution can be written as
\begin{align}
  X(r,l)=r R- l Y\,.
\end{align}

>>>>

The results for $\Delta N_{\text{eff}}$ as a function of $M_{Z^{\prime}}/g^{\prime}$ are displayed in Fig.~\ref{fig:Neff}, where it can be observed that for small ratios $M_{Z'}/g'$  the variation in the number of relativistic degrees of freedom of the new species is large ($\Delta N_{\text{eff}}\gtrsim0.3$ corresponds to decoupling temperatures $T^{\nu_R}_{\text{dec}}\lesssim2$ GeV) and vice versa.
% The bound that come from Planck~\cite{Aghanim:2018eyx} is stronger than the bound of LEP~\cite{Heeck:2014zfa}. and even LHC, see below
It follows that for the $\operatorname{U}(1)_{G}$ model the Planck+BAO upper bound at $2\sigma$ (solid gray line) demands that $M_{Z^{\prime}}/g^{\prime} \gtrsim 40$ TeV while for the $\operatorname{U}(1)_{R}$ model a more stringent lower bound applies $M_{Z^{\prime}}/g^{\prime} \gtrsim 60$ TeV (these bounds become slightly weaker once the Planck+BAO+$H_0$ combination~\cite{Aghanim:2018eyx} is considered, but in such a case the BBN bound from the primordial abundances of light elements would rule~\cite{Pitrou:2018cgg}). 
%in order to satisfy the conditions of $N_{\text{eff}}$ for all the symmetries studied here.
In other words, bearing in mind that $M_{Z'}/g'=3v_S/2$ we have that the energy scale of the $\operatorname{U}(1)_{X}$ symmetry breaking must be at least $\sim26$ TeV.      
On the other hand, it is remarkable the fact the next generation of CMB experiments~\cite{Benson:2014qhw,Abitbol:2019nhf} has the potential to entirely probe all the not-so minimal models~\cite{Abazajian:2019oqj}.  
%Prospect on $N_{\text{eff}}$~\cite{Abazajian:2019oqj}

@@

\section*{Acknowledgments}
Work supported by Sostenibilidad-UdeA and the UdeA/CODI Grant
2017-16286, and by COLCIENCIAS through the Grant 111577657253. D.R
would also like to thank the IIP-UFRN, for its support,
hospitality and the nice and stimulating atmosphere during the
Visiting Professors program where part of this work was completed.
\appendix
\section{$\operatorname{U}(1)_X$ anomaly cancellation }
We use $f$ ($f$) to denote the general $\operatorname{U}(1)_X$ generation-independent charge assignments of the field $f_R$ ($F_L$). The three linear anomalies in $\operatorname{U(1)_X}$~\cite{Campos:2017dgc}
\begin{align}
  \label{eq:anolin}
  \left[\mathrm{SU}(3)_{C}\right]^{2} \mathrm{U}(1)_{X} :& &
[3 u+3 d]-[3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{SU}(2)_{L}\right]^{2} \mathrm{U}(1)_{X} :&&
-[2 l+3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{U}(1)_{Y}\right]^{2} \mathrm{U}(1)_{X} :&&
{ \left[(-2)^{2} e+3\left(\tfrac{4}{3}\right)^{2} u+3\left(-\tfrac{2}{3}\right)^{2} d\right]-\left[2(-1)^{2} l+3 \cdot 2\left(\tfrac{1}{3}\right)^{2} q\right]}=&
 0,
  \end{align}
allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $N$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \left[\text{Grav}\right]^{2} \mathrm{U}(1)_{X}:\ \sum_{\alpha=1}^N n_{\alpha}+3 (e-2l)=&0\,, &  \left[\mathrm{U}(1)_{X}\right]^{3}:\ \sum_{\alpha=1}^N n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
We choose the solutions with $r\equiv e-2l$, such that
\begin{align}
  \label{eq:anolam}
  \sum_{\alpha=1}^{N} n_{\alpha}=&-3 r\,,   & \sum_{\alpha=1}^{N} n_{\alpha}^3=&-3 r^3\,.
\end{align}
The full set of anomaly free standard model $X$-charges in terms of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc} that we choose as $l$ and $r$, is just
\begin{align}
  u=&-r-\frac{4l}{3}\,,&d=&r+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&r+2l\,,&h=&-r-l\,.
  \label{Eq:SMCharges}
\end{align}
where the condition in the charged lepton Yukawa couplings have been used to fix $h$, and is automatically consistent with the conditions in the quark Yukawa couplings. By setting $l=0$ in the previous equations, we can define the Abelian symmetry in which only the right-handed charged fermions have non-vanishing $X$-charges as $\operatorname{U}(1)_R$. Then the general anomaly free two-parameter solution can be written as
\begin{align}
  X(r,l)=r R- l Y\,.
\end{align}
